---
title: "htmltools: from R to HTML"
author: "David Granjon"
date: "05/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(htmltools)
library(magrittr)
library(purrr)
```

## Exercise 1: tags structure (5 minutes)
Consider the following shiny tag:

```{r}
myTag <- a(
  class = "btn btn-large",
  type = "button",
  span(class = "child1", id = "super-span", 1),
  span(class = "child2", 2)
)

myTag
```

1. Inspect its structure. Hint: you may use `str`.
2. Access its class using `tagGetAttribute` and another method of your choice.
3. Modify the first child class to `custom class`.

## Exercise 2: modifiying tags (5 minutes)
Let us consider the following tag:

```{r}
temp <- div("Hello World")
```

(You may chain functions with `%>%`)

1. Replace its unique child by `a(href = "http://www.google.com", "click me!")`. Hint: `tagSetChildren` is your friend.
2. Add 10 other `span`. Hint: tags may be programmatically generated with `lapply` or `purrr::map`.