---
title: "HTML dependencies"
author: "David Granjon"
date: "06/06/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(htmltools)
library(shiny)
```

## Exercise 1: html dependencies (10 minutes)
As shown earlier, Shiny is built on top of the Bootstrap 3 framework. We did check with the HTML inspector. `{htmltools}` has utils function to explore element dependencies like `findDependencies`. let's try on a simple `fluidPage` element.

```{r}
findDependencies(fluidPage())
# Adding some icon.
findDependencies(
  fluidPage(
    icon("trash")
  )
)
```

Adding an extra icon component brings another dependency!


We would like to create a new Bootstrap 4 card component. The code is already provided.

```{r}
# we create the card function before
my_card <- function(...) {
  withTags(
    div(
      class = "card border-success mb-3",
      div(class = "card-header bg-transparent border-success"),
      div(
        class = "card-body text-success",
        h3(class = "card-title", "title"),
        p(class = "card-text", ...)
      ),
      div(class = "card-footer bg-transparent border-success", "footer")
    )
  )
}
```


1. This card does not render well, as you can see below:

```{r}
ui <- fluidPage(my_card("test"))
server <- function(input, output, session) {}
shinyApp(ui, server)
```

Try to explain why.

2. Given that we may found the Bootstrap 4 CSS at https://maxcdn.bootstrapcdn.com/bootstrap/4.0.0/css/bootstrap.min.css, create an HTML dependency containing this reference. You may start with the template below.

```{r}
card_deps <- htmlDependency(
  name = "",
  version = "",
  src = c(href = ""),
  stylesheet = ""
)
```

3. Attach it to the card. Hint: apply the `tagList(tag, deps)` pattern.

```{r}
# we create the card function before
my_card <- function(...) {
  tag <- withTags(
    div(
      class = "card border-success mb-3",
      div(class = "card-header bg-transparent border-success"),
      div(
        class = "card-body text-success",
        h3(class = "card-title", "title"),
        p(class = "card-text", ...)
      ),
      div(class = "card-footer bg-transparent border-success", "footer")
    )
  )
  
  # bind the dependency below. Complete the code
  tagList()
}
```

4 Create a Shiny app to check your work. Hint: typing `shinyapp` in the RStudio IDE triggers a snippet that automatically created a super simple shiny app skeleton. 

